Diverse Genotype-by-Weather Interactions in Switchgrass

### 2024-05-24 manuscript/revision2/revision2.qmd
revision: Address Tom's comment about sign-changing GxE vs antagonistic pleiotropy

  The timing of vegetative and reproductive growth in plants ("phenological timings") depends on genetic effects, environmental (e.g., weather) cues, and possibly their interaction. Quantifying these GxWeather interactions can aid prediction and manipulation of phenological timings. Here, we map GxWeather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of cloned individuals from these populations at eight sites across the central United States. Some genetic effects covary with  interpretable weather-based cues. For example, in the Gulf population, 65% of genetic effects on the timing of vegetative growth covary with daylength 14 days prior to green-up date, and 33% of genetic effects on the timing of flowering covary with cumulative rainfall in the seven days prior to flowering. However, most variation in genetic effects across sites is site-specific and can not be attributed to variation in weather variables alone. We demonstrate that we can identify genetic variation with GxWeather and assign these loci to specific weather-based cues or other patterns. Breeding for particular alleles at the GxWeather loci we identified could change flowering responsiveness in a photoperiod and rainfall-specific way. More broadly, our approach proposes a refined, powerful characterization of genotype-by-environment interactions in any species phenotyped in multiple environments.

@@

keywords:  [allele-by-environment effect variation, antagonistic pleiotropy, photoperiod, cumulative rainfall, genetic variation]

@@

Flowering time, or the transition from vegetative to reproductive growth, is a common subject of GxE research [@blackman2013interacting; @henry2014transitions; @aagren2017adaptive; @brachi2010linkage; @dittmar2014flowering; @li2018genomic; @romero2017study], a key output of selection driving adaptation to local environments [@andres2012genetic; @wadgymar2017identifying; @blumel2015flowering], and a selection target for crop improvement to adapt crops to local or future environments @jung2009flowering. Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in several cereal species [@turner2005pseudo; @faure2012mutation; @hung2012zmcct; @zakhrabekova2012induced; @yang2013oself3] and other crops [@pin2012multifaceted; @weller2019parallel]. Recent statistical advances in studying phenological GxE have involved determining critical environmental indices before the phenological event occurs, such as photothermal time within a critical growth window @li2018genomic. However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE [@Weine2023.06.21.545998]. Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or sign-changing GxE, at the level of individual loci [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous empirical work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying]. However, this work has been limited by a known statistical bias that reduced detection of genetic effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary]. Thus, despite substantial interest in the frequencies of various forms of GxE, the prevalence of antagonistic pleiotropy relative to other forms of GxE remains unknown.

@@

Here, we estimate GxWeather for two phenological timings in switchgrass by assigning patterns of genetic effects on phenology across gardens to many patterns of weather covariance at these gardens. To do this, we phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative and reproductive development at eight common garden locations spanning 17 degrees of latitude. These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters. We define multiple ways phenological timings might covary with weather (Table 1) and additional ways phenological timings might vary by site (SI Appendix, Table S1, Section SX), then jointly re-estimate genetic effects on these timings at all eight sites using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (SI Appendix, Section SX) @urbut2019flexible. We use the Bayesian framework *mash* (multivariate adaptive shrinkage) developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight sites. *mash* allows us to identify and specify multiple covariance structures among genetic effect estimates across sites, including structures that represent covariance in weather variables of interest. Importantly, this method circumvents statistical biases in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041. To confirm our genetic mapping of GxWeather, we compare our posterior genetic effect estimates from *mash* to mapping results from an outbred pseudo-F2 cross grown at the same sites. Our analyses allow us to describe the weather cues and genetic variation affecting phenology in two divergent natural populations of switchgrass.

@@

We next looked for evidence of GxWeather by jointly re-estimating genetic effects across all eight common gardens using the *mash* model, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Datasets 1-6). We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S2). For example, the more similar that the amount of rainfall prior to flowering is for genotypes A and B between gardens in MI and MO, the nearer these covariances will be to one. If rainfall amounts are uncorrelated, covariance will be near zero; if there is rank changing (e.g. high rainfall before flowering of A at MO but not MI, and high rainfall before flowering for B at MI but not MO), this covariance will be near negative one. We specified multiple covariance structures for the canonical, data-driven, and GxWeather categories, each of which represented a specific pattern of covariance in genetic effect size estimates that could be included in the model. The phenotypic correlations for the timing of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf subpopulation and Both subpopulations (@fig-map B). If these phenotypic correlations have a genetic component, they could be partially or completely captured by the covariance structures specified in the *mash* model.

The GxWeather covariance structures allow hypothesis testing of specific weather variables as cues for the start of vegetative and reproductive growth. Say that a SNP in FLC controls flowering in a photoperiod-dependent manner. In that case, the joint estimate of effects for that SNP could have a high mixture proportion, or mass, on a covariance matrix created using a photoperiod-based environmental cue, such as day length at some interval prior to flowering. In our data, we would infer that the effect of that SNP on flowering was caused by a response to the environmental cue used to construct the GxWeather covariance structure with the largest mass. Overall, loadings of genetic effects on these GxWeather matrices in the model, along with the posterior effect size estimates, provide information on genome-wide patterns of SNP-environment interaction.

@@

For flowering date, distinct GxWeather covariance structures captured covariance in effect sizes for SNPs in the Gulf and Midwest subpopulations. 33% of SNP effects on flowering in the Gulf subpopulation covaried with cumulative rainfall in the seven days prior to flowering (@fig-covar D). 22.6% of SNP effects on flowering in the Midwest subpopulation covaried with day length change in the two days prior to flowering (@fig-covar D), which had negative covariances between Texas and North gardens. Neither of these covariance matrices significantly improved the log-likelihood of the *mash* model of Both subpopulations, and no GxWeather covariance matrices had non-zero mass in models of Both subpopulations (@fig-covar E). In five of the six *mash* models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (@fig-covar C,E); in these five models, the majority of this mass was on various canonical covariance matrices. These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and garden-specific effects present only at one garden.

We next characterized the pairwise patterns of effects where we were confident in the sign of the effect at both gardens. If the overall pattern of phenotypic expression differs between a pair of gardens, this difference may be comprised of SNP effects that differ in sign between these gardens (effects with antagonistic pleiotropy), SNP effects that differ in magnitude (effects that are large in one garden or region and smaller in others), and SNP effects that are indistinguishable in the two gardens (@fig-effects A). We used the local false sign rate (lfsr), an analogue of the lfdr that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance. We required lfsr significance (p \< 0.05) in both gardens to include effects. This means that our tests for antagonistic pleiotropy, or a sign change between conditions, carry an equal statistical burden to those for effects with the same sign.

For green-up date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, between pairs of Texas and North gardens (@fig-effects B; SI Appendix, Fig SX). 78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (@fig-effects B). The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects most often differed in magnitude between gardens within and between the regions (@fig-effects B).

For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited antagonistic pleiotropy, or a difference in effect sign, within or between regions (@fig-effects B). More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; @fig-effects B), while the majority of effects had no GxE. The Midwest population had relatively few significant effects for flowering, but a large proportion of these differed in sign between Texas and North regions (42.7%) or within the North region (65.4%). Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (@fig-effects B). Most differences in sign were between TX1, the southernmost garden, and all other gardens (SI Appendix, Fig SX). Similarly, more effects that differed in magnitude included gardens in the Texas region (52.3-55.9%), and most effect pairs in the North region were not distinguishable (91.5%).

@@

We sought additional experimental support for our joint re-estimates of SNP effects using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (@fig-qtl A,B). We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in *mash*, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped one or more homologs from rice or *A. thaliana* with functionally validated roles in flowering (SI Appendix, Dataset 7). All flowering and green-up QTL intervals contained at least one SNP significant in at least one *mash* run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter (SI Appendix, Dataset 8). We also looked for enrichments of *mash* SNPs in the 1% tail of significance (the '*mash* 1% tail') within each QTL interval. At the 5% level, three QTL had enrichments of SNPs in the *mash* 1% tail. Overall, there were five significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the *mash* 1% tail in the QTL intervals. Thus, we were able to experimentally support some of our re-estimates of SNP effects with a QTL mapping experiment using a separate mapping population.

@@

As the climate and the natural environment change, it is increasingly critical to understand how patterns of plant-environment interactions will change in response. To do this, we must understand the current patterns of trait covariation across environments, the genetic underpinnings of these patterns, and the cases where this covariation can be altered. Here, we demonstrate that we can associate multiple patterns of GxWeather with specific genomic regions using a switchgrass diversity panel grown at eight common gardens. We can assign genetic effects to both GxWeather patterns with interpretable weather-based cues, and to agnostic, site-based patterns. We use this approach to study GxWeather for the timings of vegetative and reproductive development in the deeply genetically diverged Gulf and Midwest subpopulations of switchgrass.

Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed substantial antagonistic pleiotropy in effects between the Texas and North gardens (@fig-effects B). This result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic], and is the first experimental work using GWAS across common gardens to find antagonistic pleiotropy to be common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].

Our analysis of the timing of flowering showed that the Gulf and Midwest subpopulations have distinct GxWeather: flowering timing in the Midwest subpopulation has photoperiod-related variation, in that flowering timing covaries with a day length change signal two days before flowering occurs. In contrast, the Gulf subpopulation does not have variation in flowering based on a photoperiod cue. Instead, the Gulf subpopulation has variation in flowering that covaries with the rainfall that occurs in the week prior to flowering. Three genomic regions affecting flowering that we re-estimated across all eight sites were also supported by QTL from an independent mapping population at these sites (@fig-qtl B).

@@

Last, we characterized the overall patterns of antagonistic pleiotropy in the set of SNPs where there was pairwise significance of effects at pairs of gardens. To do this, we used the 'get_GxE' function of the switchgrassGWAS R package. First, this determines the set of SNPs with evidence of significant effects in both conditions for all pairs of conditions using local false sign rates (lfsr) as the significance criteria. Second, to determine antagonistic pleiotropy, this function determines if effects significant in both conditions are of opposite sign.

Using the lfsr rather than the local false discovery rate (lfdr) is a critical change in our ability to detect antagonistic pleiotropy. The lfdr, like other measures of FDR, focuses on if we have enough evidence to reject the null hypothesis that an effect j is 0, or that there is a significant effect. Previous studies of antagonistic pleiotropy (e.g. [@lowry2019qtl]) have used the lfdr or equivalent statistical tests to detect antagonistic pleiotropy. These tests were conservative, in that they required two non-zero effects of different signs, while tests for differential sensitivity required only one non-zero effect. This previous work recognized that this testing bias could lead to undercounting occurrences of antagonistic pleiotropy [@anderson2011evolutionary; @des2013genotype], and sought to reduce it by permutation [@anderson2013genetic]. However, using the lfsr to test for antagonistic pleiotropy does not undercount occurrences of antagonistic pleiotropy, as this statistic answers a fundamentally different question. For each effect j, the lfsr~j~ is defined as the probability that we make an error in the sign of effect j if we were forced to declare the effect positive or negative @10.1093/biostatistics/kxw041. Thus, rather than asking "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - the local false sign rate answers a more meaningful question: Can we be confident in the sign of this effect?

In addition, the get_GxE function also sets an arbitrary threshold to count an effect as changing in magnitude between environments, commonly known as differential sensitivity or a change in amplitude of the effect. For differential sensitivity, this function determines if effects significant in both conditions are of the same sign and of a magnitude (not tested for significance) that differs by a factor of 0.4 or more. The remaining effects that are significant in both conditions have the same effect sign and similar effect magnitudes and we denote these effects as having no GxE. The distinction between effects with different magnitudes is arbitrary but useful to fully characterize how effects vary across environments to ultimately influence phenotypes. Our use of the lfsr to determine significance and our specification that SNP effects must be significant in both conditions to be included means that our tests for antagonistic pleiotropy carry an equal statistical burden to those measuring differential sensitivity and effects without GxE.

>>>>

  The timing of vegetative and reproductive growth in plants ("phenological timings") depends on genetic effects, environmental (e.g., weather) cues, and possibly their interaction. Quantifying these GxWeather interactions can aid prediction and manipulation of phenological timings. Here, we map GxWeather effects on phenological timings in two highly divergent switchgrass (*Panicum virgatum*) populations using repeated plantings of cloned individuals from these populations at eight sites across the central United States. Some genetic effects covary with  interpretable weather-based cues. For example, in the Gulf population, 65% of genetic effects on the timing of vegetative growth covary with daylength 14 days prior to green-up date, and 33% of genetic effects on the timing of flowering covary with cumulative rainfall in the seven days prior to flowering. However, most variation in genetic effects is site-specific and can not be attributed to variation in the studied weather variables. We demonstrate that we can identify genetic variation with GxWeather and assign these loci to specific weather-based cues or other patterns. Breeding for particular alleles at the GxWeather loci we identified could change flowering responsiveness in a photoperiod or rainfall-specific way. More broadly, our approach proposes a refined, powerful characterization of genotype-by-environment interactions in any species phenotyped in multiple environments.

@@

keywords:  [allele-by-environment effect variation, rank-changing genotype-by-environment interaction, photoperiod, cumulative rainfall, genetic variation]

@@

Flowering time, or the transition from vegetative to reproductive growth, is a common subject of GxE research [@blackman2013interacting; @henry2014transitions; @aagren2017adaptive; @brachi2010linkage; @dittmar2014flowering; @li2018genomic; @romero2017study], a key output of selection driving adaptation to local environments [@andres2012genetic; @wadgymar2017identifying; @blumel2015flowering], and a selection target for crop improvement to adapt crops to local or future environments @jung2009flowering. Changing flowering responsiveness to photoperiod cues has allowed geographic range expansion and increased yields in several cereal species [@turner2005pseudo; @faure2012mutation; @hung2012zmcct; @zakhrabekova2012induced; @yang2013oself3] and other crops [@pin2012multifaceted; @weller2019parallel]. Recent statistical advances in studying phenological GxE have involved determining critical environmental indices before the phenological event occurs, such as photothermal time within a critical growth window @li2018genomic. However, most studies of flowering GxE focus on finding a single, best fitting form of genotype-environment covariance, despite the key expectation that different genetic subpopulations, and even different genomic regions, have likely evolved distinct patterns of GxE [@Weine2023.06.21.545998]. Additionally, despite theoretical predictions that local adaptation should involve antagonistic pleiotropy, or alleles with effects with opposing fitness outcomes [@levene1953genetic; @felsenstein1976theoretical; @kawecki2004conceptual; @hedrick1986genetic], previous empirical work has found limited evidence of antagonistic pleiotropy [@des2013genotype; @wadgymar2017identifying]. However, this work has been limited by a known statistical bias that reduced detection of genetic effects that differ in sign [@des2013genotype; @anderson2013genetic; @anderson2011evolutionary]. Thus, despite substantial interest in the frequencies of various forms of GxE, particularly sign-changing GxE, relative to other forms of GxE remains unknown.

@@

Here, we estimate GxWeather for two phenological timings in switchgrass by assigning patterns of genetic effects on phenology for cloned individuals grown at eight common gardens to many patterns of weather covariance at these gardens. To do this, we phenotype a diversity panel of hundreds of switchgrass genotypes from the Midwest and Gulf subpopulations for the timing of vegetative and reproductive development at eight common garden locations spanning 17 degrees of latitude. These gardens cover the majority of the latitudinal and climatic range of switchgrass and capture the most comprehensive picture to date of the environmental variation this species encounters. We define multiple ways phenological timings might covary with weather (Table 1) and additional ways phenological timings might vary by garden (SI Appendix, Section S1), then jointly re-estimate genetic effects on these timings at all eight common gardens using the set of these covariance matrices that significantly improved the modeled log-likelihood when included (SI Appendix, Section S2) @urbut2019flexible. We use the Bayesian framework *mash* (multivariate adaptive shrinkage) developed by @urbut2019flexible, to refine effect size estimates from genome-wide association (GWAS) conducted on individuals at each of the eight gardens. *mash* allows us to identify and specify multiple covariance structures among genetic effect estimates across sites, including structures that represent covariance in weather variables of interest. Importantly, this method circumvents statistical biases in detecting genetic effects with the same or opposite signs @10.1093/biostatistics/kxw041. To confirm our genetic mapping of GxWeather, we compare our significant posterior genetic effects locations from *mash* to mapping results from an outbred mapping population grown at the same sites. Our analyses allow us to describe the weather cues and genetic variation affecting phenology in two divergent natural populations of switchgrass across the species' latitudinal range.

@@

We next looked for evidence of GxWeather by jointly re-estimating genetic effects of SNPs across all eight common gardens using the *mash* model, which can flexibly capture many modes of covariance across gardens (@fig-map D; SI Appendix, Datasets 1-6). We specified three qualitative categories of covariance structures across gardens: "canonical" covariance, with simple patterns of effect size covariance introduced in the initial *mash* manuscript; "data-driven" covariances derived from common patterns of SNP effects observed in the data, and "GxWeather" covariance, estimated from the covariance of empirical weather patterns at each garden at specific times before the phenological event (Table 1; SI Appendix, Section S2). For example, the more similar that the amount of rainfall prior to flowering is for genotypes A and B between gardens in MI and MO, the nearer these covariances will be to one. If rainfall amounts are uncorrelated, covariance will be near zero; if there is rank changing (e.g. high rainfall before flowering of A at MO but not MI, and high rainfall before flowering for B at MI but not MO), this covariance will be near negative one. We specified multiple covariance structures for the canonical, data-driven, and GxWeather categories, each of which represented a specific pattern of covariance in genetic effect size estimates that could be included in the model. The phenotypic correlations for the timing of vegetative growth had moderate negative correlations between the Texas and North gardens, particularly in the Gulf subpopulation and Both subpopulations (@fig-map B). If these phenotypic correlations have a genetic component, they could be partially or completely captured by the covariance structures specified in the *mash* model.

The GxWeather covariance structures allow hypothesis testing of specific weather variables as cues for the start of vegetative and reproductive growth. Say that a SNP in the gene *Flowering Locus C* (FLC), a well-known flowering time regulator, controls flowering in a photoperiod-dependent manner. In that case, the joint estimate of effects for that SNP could have a high mixture proportion, or mass, on a covariance matrix created using a photoperiod-based environmental cue, such as day length at some interval prior to flowering. In our data, we would infer that the effect of that SNP on flowering was caused by a response to the environmental cue used to construct the GxWeather covariance structure with the largest mass. Overall, loadings of genetic effects on these GxWeather matrices in the model, along with the posterior effect size estimates, provide information on genome-wide patterns of SNP-environment interaction.

@@

For flowering date, distinct GxWeather covariance structures captured covariance in effect sizes for SNPs in the Gulf and Midwest subpopulations. 33% of SNP effects on flowering in the Gulf subpopulation covaried with cumulative rainfall in the seven days prior to flowering (@fig-covar D). 22.6% of SNP effects on flowering in the Midwest subpopulation covaried with day length change in the two days prior to flowering (@fig-covar D), which had negative covariances between Texas and North gardens. Neither of these covariance matrices significantly improved the log-likelihood of the *mash* model of Both subpopulations, and no GxWeather covariance matrices had non-zero mass in models of Both subpopulations (@fig-covar E). In five of the six *mash* models of strong effects, the GxWeather covariance matrices captured a minority of the posterior weights of the strong effects (@fig-covar C,E); in these five models, the majority of this mass was on various canonical covariance matrices. These matrices included simple heterozygosity, with intermediate, positive covariances between all gardens, and single effect matrices with garden-specific effects (SI Appendix, Section S1).

We next characterized the pairwise patterns of effects where we were confident in the sign of the effect at both gardens. If the overall pattern of phenotypic expression differs between a pair of gardens, this difference may be comprised of SNP effects that differ in sign between these gardens (effects with rank-changing GxE between gardens), SNP effects that differ in magnitude (effects that are large in one garden or region and smaller in others), and SNP effects that are indistinguishable in the two gardens (@fig-effects A). We used the local false sign rate (lfsr), an analogue of the lfdr that establishes confidence in the effect sign, not the effect's difference from zero, to determine significance. We required lfsr significance (p \< 0.05) in both gardens to include effects. This means that our tests for a sign change between gardens carry an equal statistical burden to those for effects with the same sign.

For green-up date for the Gulf subpopulation, hundreds to thousands of pairwise effects exhibited a difference in effect sign, between pairs of Texas and North gardens (@fig-effects B; SI Appendix, Fig SX). 78.7% of pairwise comparisons between North and Texas gardens had a difference in sign, while only 28.6% and 0.2% of North-North or Texas-Texas comparisons had a difference in sign, respectively (@fig-effects B). The majority of pairwise effects for greenup for the Midwest (\>55%) and Both (\>85%) subpopulations were the same sign, and effects most often differed in magnitude between gardens within and between the regions (@fig-effects B).

For flowering date for the Gulf subpopulation, less than 2% of pairwise effects exhibited a difference in effect sign within or between regions (@fig-effects B). More effects differed in magnitude between the Texas and North regions than within these regions (42.7% vs \<20%; @fig-effects B), while the majority of effects had no GxE. The Midwest population had relatively few significant effects for flowering, but a large proportion of these differed in sign between Texas and North regions (42.7%) or within the North region (65.4%). Finally, in Both subpopulations, less than 20% of pairwise effects differed in sign (@fig-effects B). Most differences in sign were between TX1, the southernmost garden, and all other gardens (SI Appendix, Fig SX). Similarly, more effects that differed in magnitude included gardens in the Texas region (52.3-55.9%), and most effect pairs in the North region were not distinguishable (91.5%).

@@

We sought additional experimental support the occurrence and location of our significant SNP effect re-estimates using an independent pseudo-F2 mapping population created from Gulf & Midwest individuals and grown at the same sites (@fig-qtl A,B). We conducted quantitative trait loci (QTL) mapping of flowering as functions of five environmental cues that we also used as covariance matrices in *mash*, and identified eight QTL for flowering date, eight QTL for flowering as a function of day length change two days prior, one QTL for the start of vegetative growth, and two QTL for vegetative growth as a function of daylength change one day prior, all of which showed QTL by environment interactions (SI Appendix, Fig. S3). All QTL for flowering overlapped one or more homologs from rice or *A. thaliana* with functionally validated roles in flowering (SI Appendix, Dataset 7). All flowering and green-up QTL intervals contained at least one SNP significant in at least one *mash* run at a log10-transformed Bayes Factor \> 2, or in the 1% tail of significance, whichever was stricter (SI Appendix, Dataset 8). We also looked for enrichments of *mash* SNPs in the 1% tail of significance (the '*mash* 1% tail') within each QTL interval. At the 5% level, three QTL had enrichments of SNPs in the *mash* 1% tail. Overall, there were five significant enrichments (p \< 0.05, hypergeometric test) of SNPs in the *mash* 1% tail in the QTL intervals. Thus, we were able to experimentally support some of our re-estimates of SNP effects with a QTL mapping experiment using a separate mapping population.

@@

As the climate and the natural environment change, it is increasingly critical to understand how patterns of plant-environment interactions will change in response. To do this, we must understand the current patterns of trait covariation across environments, the genetic underpinnings of these patterns, and the cases where this covariation can be altered. Here, we demonstrate that we can associate multiple patterns of GxWeather with specific genomic regions using a switchgrass diversity panel grown at eight common gardens. We can assign genetic effects to both GxWeather patterns with interpretable weather-based cues, and to unmeasured, site-based patterns. We use this approach to study GxWeather for the timings of vegetative and reproductive development in the deeply genetically diverged Gulf and Midwest subpopulations of switchgrass.

Our analysis of the timing of vegetative growth in the Gulf and in Both subpopulations revealed many alleles with sign changes in their effects between the Texas and North gardens (@fig-effects B). As phenological timings are major components of plant fitness, this result supports theoretical models that local adaptation should involve antagonistic pleiotropy at the level of individual loci [@kawecki2004conceptual; @felsenstein1976theoretical; @levene1953genetic; @hedrick1986genetic]; it expands the genetics of local adaptation research to more than two field sites and to a wider range of genetic variation by using effects from GWAS in common gardens located throughout the species' range to determine that rank-changing GxE for phenological timings is common in small genomic regions [@wadgymar2017identifying; @savolainen2013ecological; @lowry2019qtl].

Our analysis of the timing of flowering showed that the Gulf and Midwest subpopulations have distinct GxWeather: flowering timing in the Midwest subpopulation has photoperiod-related genetic variation, in that flowering timing covaries with a day length change signal two days before flowering occurs. In contrast, the Gulf subpopulation does not have genetic variation in flowering that covaries with a photoperiod cue. Instead, the Gulf subpopulation has genetic variation in flowering that covaries with the rainfall that occurs in the week prior to flowering. Three genomic regions affecting flowering that we re-estimated across all eight sites were also supported by QTL from an independent mapping population at these sites (@fig-qtl B).

@@

Last, we characterized the overall patterns of GxE in the set of SNPs where there was pairwise significance of effects at pairs of gardens. To do this, we used the 'get_GxE' function of the switchgrassGWAS R package. First, this determines the set of SNPs with evidence of significant effects in both conditions for all pairs of conditions using local false sign rates (lfsr) as the significance criteria. Then, this function determines if effects significant in both conditions are of opposite sign.

Using the lfsr rather than the local false discovery rate (lfdr) is a critical change in our ability to detect alleles directly contributing to rank-changing GxE between environments. The lfdr, like other measures of FDR, focuses on if we have enough evidence to reject the null hypothesis that an effect j is 0, or that there is a significant effect. Previous studies of antagonistic pleiotropy (e.g. [@lowry2019qtl]) have used the lfdr or equivalent statistical tests to detect antagonistic pleiotropy. These tests were conservative, in that they required two non-zero effects of different signs, while tests for differential sensitivity required only one non-zero effect. This previous work recognized that this testing bias could lead to undercounting occurrences of antagonistic pleiotropy [@anderson2011evolutionary; @des2013genotype], and sought to reduce it by permutation [@anderson2013genetic]. However, using the lfsr to test for allelic effects that differ in sign does not undercount these occurences, as this statistic answers a fundamentally different question. For each effect j, the lfsr~j~ is defined as the probability that we make an error in the sign of effect j if we were forced to declare the effect positive or negative @10.1093/biostatistics/kxw041. Thus, rather than asking "Are these two effects different?" - as we reasonably expect two effects to be, even if this difference cannot be measured - the local false sign rate answers a more meaningful question: Can we be confident in the sign of this effect?

In addition, the get_GxE function also sets an arbitrary threshold to count an effect as changing in magnitude between environments, commonly known as differential sensitivity or a change in amplitude of the effect. For differential sensitivity, this function determines if effects significant in both conditions are of the same sign and of a magnitude (not tested for significance) that differs by a factor of 0.4 or more. The remaining effects that are significant in both conditions have the same effect sign and similar effect magnitudes and we denote these effects as having no GxE. The distinction between effects with different magnitudes is arbitrary but useful to fully characterize how effects vary across environments to ultimately influence phenotypes. Our use of the lfsr to determine significance and our specification that SNP effects must be significant in both conditions to be included means that our tests for alleles with rank-changing GxE carry an equal statistical burden to those measuring differential sensitivity and effects without GxE.
